---
title: "Ridge and Lasso"
author: "Andrew Costa"
date: "07/06/2022"
output: 
  html_document: default
  word_document: default

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exploration of Ridge and Lasso Regression: Wine dataset

The wine dataset is one particulary widespread given its usability. It is an organised and clean dataset, with no missing values and mostly linear variables.

The goal of this part of the project will be to firstly reduce the number of variables in the model in order to have a more comprehensible model output. The second part will be to divide the quality of wine into binary categories -- good and bad -- in order to then classify the wines.

The project will be divided into the following steps:
  1. Exploratory Data Analysis
  2. Data Cleaning
  3. Feature Engineering 
      - Best subset selection
      - Forward subset selection
      - Backward subset selection
  4. Model Creation
      - Initialise Logistic Regression
      - Ridge Regression
      - Lasso
  5. Model Evaluation


## 1. Import any dependencies needed
```{r library}
library(readxl)
library(readr)
library(dplyr)
library(magrittr)
library(ggpubr)
library(car)
library(rsample)
library(dplyr)
library(bestglm)
library(dplyr)
setwd("/Users/andrewcosta/Desktop/Milan/Statistical_Learning/Data_Silence")
source("utils.R")
source("packages.R")
```

## 2. Data Preprocessing

```{r concat, echo=FALSE}
getwd()
rw = read.table("/Users/andrewcosta/Desktop/Milan/Statistical_Learning/Data_Silence/winequality-red.csv",sep=",",header=T)
ww = read.table("/Users/andrewcosta/Desktop/Milan/Statistical_Learning/Data_Silence/archive/winequality-white.csv",sep=";",header=T)
source("utils.R")

rw['type'] <- c("red")
ww['type'] <- c("white")
wdb <- rbind(ww,rw)
wdb<-as.data.frame(wdb)
summary(wdb)

####Data fixing####
wdb <- wdb %>%
  mutate(
    quality = ifelse(quality<=5, "low", "high"),
    quality = factor(quality)
  ) %>%
  mutate(across(where(is.character), as.factor))
wdb <- wdb[!duplicated(wdb),]

##PREPROCESSING####
###Train/test split####
set.seed(42)
split <- initial_split(wdb, strata=type, prop = .8)
splits <- initial_split(training(split), strata=type, prop= .7)
folds <- vfold_cv(training(splits))
qty_split <- initial_split(wdb %>% dplyr::select(-quality), strata = type)
```



## 3. Baseline Model Initiation

```{r}
# initialise baseline recipe
###Baseline recipe####  
rcp_spec <- recipe(quality ~ ., data = training(splits))%>%
  step_dummy(type)
rcp_spec %>% prep() %>% juice() %>% glimpse()
```


```{r}
# initialise baseline model
model_logistic <- glm(quality ~., data = training(splits), family = binomial)
# view model coefficients
coef(model_logistic)

```

## 3. Parameter Selection
### Best Subset Selection

```{r}
wdb.bglm <- wdb[,c(1:12)]
wdb.bglm <- wdb.bglm %>%
  mutate(
    quality = ifelse(quality=="low", 0, 1),
    quality = factor(quality)
  )
wdb.bglm = rename(wdb.bglm, y=quality)
Xy <- wdb.bglm

model_best_subset_selection <- bestglm(Xy, IC = "AIC", family=binomial)
model_best_subset_selection
summary(model_best_subset_selection)

model_best_subset_selection$Subsets

```

### Forward Subset Selection

```{r}
library(MASS)
model_forward_selection <- model_logistic %>% stepAIC(trace = FALSE, direction = "forward")
coef(model_forward_selection)
```

### Backwards Subset Selection

```{r}
model_backward_selection <- model_logistic %>% stepAIC(trace = FALSE, direction = "backward")
coef(model_backward_selection)

```


## 4. Model Initiation: Ridge, LASSO, ElasticNet

```{r}
# ridge
ridge_model <-  logistic_reg( #RIDGE
  mode = "classification",
  penalty = tune(), #to be tuned!
  mixture = 0
) %>%
  set_engine("glmnet")

# LASSO
lasso_model <-  logistic_reg( #LASSO
  mode = "classification",
  penalty = tune(), #to be tuned!
  mixture = 1
) %>%
  set_engine("glmnet")

# MIXED
elastic_net_model <-  logistic_reg( #ELASTIC NET (w/ mixture of norms to be tuned)
  mode = "classification",
  penalty = tune(), #to be tuned!
  mixture = tune() #to be tuned!
) %>%
  set_engine("glmnet")

####* Workflows ----------------------------------------------------------------
workflow_ridge <- workflow() %>% #RIDGE
  add_model(ridge_model) %>%
  add_recipe(rcp_spec)

workflow_lasso <- workflow() %>% #LASSO
  add_model(lasso_model) %>%
  add_recipe(rcp_spec)

workflow_elastic_net <- workflow() %>% #ELASTIC NET
  add_model(elastic_net_model) %>%
  add_recipe(rcp_spec)
```

## 5. Grid Search

```{r}
# Grid for only the penalty (to be used for lasso and ridge).
set.seed(42)
model_grid_penalty <- grid_regular( 
  penalty(),
  levels = 10
)
model_grid_penalty
model_grid_penalty %>% map(unique)

# Grid for both the penalty and the mixture (to be used for elastic net). Creates 
model_grid_penalty_mixture <- grid_regular(
  penalty(),
  mixture(),
  levels = 10
)

model_grid_penalty_mixture
model_grid_penalty_mixture %>% map(unique)
```


## 6. Tuning Parameters

```{r}
####* Tuning -------------------------------------------------------------------
set.seed(42)
# Tuning of the hyperparameter (penalty) for the Ridge
model_result_ridge <- workflow_ridge %>% 
  tune_grid(
    resamples = folds,
    grid = model_grid_penalty
  )

# Tuning of the hyperparameter (penalty) for the Lasso
model_result_lasso <- workflow_lasso %>% 
  tune_grid(
    resamples = folds,
    grid = model_grid_penalty
  )

# Tuning of the hyperparameters (penalty and mixture) for the Elastic Net
model_result_elastic_net <- workflow_elastic_net %>% 
  tune_grid(
    resamples = folds,
    grid = model_grid_penalty_mixture
  )

```


## 7. Model Evaluation

```{r}
####* Evaluation  --------------------------------------------------------------
# Select the best values of the hyperparameter choosing the value that minimizes
# the accuracy
model_result_ridge %>% collect_metrics()
model_ridge_best <- model_result_ridge %>% select_best("accuracy")
model_ridge_best
```

```{r}
model_result_lasso %>% collect_metrics()
model_lasso_best <- model_result_lasso %>% select_best("accuracy")
model_lasso_best
```

```{r}
model_result_elastic_net %>% collect_metrics()
model_elastic_net_best <- model_result_elastic_net %>% select_best("accuracy")
model_elastic_net_best
```


```{r}
####* Re-fitting  --------------------------------------------------------------
# Refit the models on the entire training set using the best value of the hyperparameters.
# Test the performance of this model on the test set.
workflow_ridge_final <-	workflow_ridge %>%	
  finalize_workflow(model_ridge_best) %>% 
  last_fit(splits) 

workflow_lasso_final <-	workflow_lasso %>%	
  finalize_workflow(model_lasso_best) %>% 
  last_fit(splits)

workflow_elastic_net_final <-	workflow_elastic_net %>%	
  finalize_workflow(model_elastic_net_best) %>% 
  last_fit(splits)



```


### Ridge Results

```{r}
workflow_ridge_final %>%	collect_metrics()
workflow_ridge_final %>% collect_predictions()
```

### LASSO Results

```{r}
workflow_lasso_final %>%	collect_metrics()
workflow_lasso_final %>% collect_predictions()
```

### Elastic Net Results

```{r}
workflow_elastic_net_final %>%	collect_metrics()
workflow_elastic_net_final %>% collect_predictions()
```

